#Employee_Attrition with Associaiton_Rules_Mining

**library needed packages
```{r}
library(arules)
library(arulesViz)
library(dplyr)
library(tidyverse)
library(readr)
library(corrplot)
library(caret)
library(ggplot2)
library(scales)
```

**Input Data
```{r}
raw_attrition <- read_csv(file = "employee_attrition.csv")
```

**data cleaning**
*feature selection*
```{r}
#remove irrelavant columns using domain knowledge 
#remove columns with only one value
raw_attrition$EmployeeCount <-NULL
raw_attrition$Over18 <-NULL
raw_attrition$EmployeeNumber <- NULL
raw_attrition$StandardHours <- NULL
str(raw_attrition)

```

*outliers*
```{r}
#check max and min and box-plot each column for outliers
summary(raw_attrition)
#found two outliers
raw_attrition[which.max(raw_attrition$TotalWorkingYears),]
#total working years=114 is found, by domain knowledge, it might be a typo of writing 14 as 114, so replace with prediction
raw_attrition$TotalWorkingYears[which.max(raw_attrition$TotalWorkingYears)] <- 14
#another outlier
raw_attrition[which.max(raw_attrition$YearsWithCurrManager),]
#years with current manager = 219 is found, might be a typo of writing 19 as 219, therefore replace with prediction
raw_attrition$YearsWithCurrManager[which.max(raw_attrition$YearsWithCurrManager)] <- 19
#did found an outlier in distance from home as 224, but remote working is possible, so replace with the second highest
raw_attrition$DistanceFromHome[which.max(raw_attrition$DistanceFromHome)] <- max(raw_attrition$DistanceFromHome[raw_attrition$DistanceFromHome!=max(raw_attrition$DistanceFromHome,na.rm=TRUE)],na.rm=TRUE)
```

*check duplicate rows*
```{r}
isTRUE(all.equal.default(nrow(raw_attrition),nrow(raw_attrition[!duplicated(raw_attrition),])))
```
no duplicate detected

*check missing values*
the sample size is 1176, which is quite small for association rules mining. Therefore, choose to replace the NAs instead of 
removing the rows with NAs
```{r}
#detect NAs in each column
sapply(raw_attrition,function(x) sum(is.na(x)))
#replace NAs with mean
raw_attrition$DistanceFromHome[which(is.na(raw_attrition$DistanceFromHome))] <- round(mean(raw_attrition$DistanceFromHome, na.rm=TRUE))
raw_attrition$JobLevel[which(is.na(raw_attrition$JobLevel))] <- round(mean(raw_attrition$JobLevel, na.rm=TRUE))
raw_attrition$PercentSalaryHike[which(is.na(raw_attrition$PercentSalaryHike))] <- round(mean(raw_attrition$PercentSalaryHike, na.rm=TRUE))
raw_attrition$PerformanceRating[which(is.na(raw_attrition$PerformanceRating))] <- round(mean(raw_attrition$PerformanceRating, na.rm=TRUE))
raw_attrition$RelationshipSatisfaction[which(is.na(raw_attrition$RelationshipSatisfaction))] <- round(mean(raw_attrition$RelationshipSatisfaction, na.rm=TRUE))
raw_attrition$TotalWorkingYears[which(is.na(raw_attrition$TotalWorkingYears))] <- round(mean(raw_attrition$TotalWorkingYears, na.rm=TRUE))
raw_attrition$YearsSinceLastPromotion[which(is.na(raw_attrition$YearsSinceLastPromotion))] <- round(mean(raw_attrition$YearsSinceLastPromotion, na.rm=TRUE))

```
two more NAs found in OverTime and Gender columns. Since our sample size is small, I do not want to delete entire rows. But for such attributes,
it is hard to give it a reasonable prediction based on other variables of the same row. So I will have 'NoInfo' as a third level for these two attributes,
and such alternation will not make any impact on association rules mining by setting appropriate support value.
```{r}
raw_attrition$OverTime[which(is.na(raw_attrition$OverTime))] <- "NoInfo"
raw_attrition$Gender[which(is.na(raw_attrition$Gender))] <- "NoInfo"
#check if there is any NA left
sapply(raw_attrition,function(x) sum(is.na(x)))
```

*check variance of each column*
```{r}
names(raw_attrition)[nearZeroVar(raw_attrition)]
```
no column is idenfified as near zero variance
data cleaning is done, ready for descriptive analysis

*EDA*
```{r}
str(raw_attrition)
#Extract columns with numerical values
num_attrition <- raw_attrition[lapply(raw_attrition,class)=="numeric"]
#make correlation matrix
cor_matrix <- cor(num_attrition,method = "pearson")
corrplot(cor_matrix,type = "upper")
```
found some volumns are positively correlated and some other columns are negatively correlated, most columns are not directly correlated strongly positively correlated:
Age & TotalWorkingYears, JobLevel & MonthlyIncome, JobLevel & TotalWorkingYears, MonthlyIncome & TotalWorkingYears, PercentSalaryHike & PerformanceRating
YearsAtCompany & YearsInCurrentRole, YearsAtCompany & YearsWithCurrentManager, YearsInCurrentRole & YearsWithCurrentManager
By common sense and domain knowledge in human resource, these correlations are very reasonable and commonly found across industries
So the results of the correlation matrix add more reliability of the dataset used

*visulization*
```{r}
#convert columns with chr values to factors
attrition <- raw_attrition
chr_col <- c("Attrition","BusinessTravel","Department","EducationField","Gender","JobRole","MaritalStatus","OverTime")
attrition[,chr_col] <- lapply((attrition[,chr_col]),as.factor)
#bar-plot of Attrition
attri_dist <- ggplot(attrition, aes(x=Attrition,y=(..count..)/sum(..count..)),fill=Attrition)+
              geom_bar()
attri_dist             
```
most of the employees are staying, less than 20% of the total quit their job
```{r}
#barplot Gender vs Attrition
gender_attri <- ggplot(attrition)+
                geom_bar(aes(x=Gender,fill=Attrition,color=Attrition))
gender_attri
```
no obvious difference of gender in the distribution of Attrition
```{r}
#density plot of MonthlyIncome vs Attrition
Income_attri <- ggplot(attrition)+
                geom_density(aes(x=MonthlyIncome,fill=Attrition),alpha=0.8)
Income_attri  
```
many employees who left the company had low imcome, interestingly, the distribution of Attrition=Yes has two local peaks at Income=10000 and 20000 
```{r}
#box plot monthlyIncome vs JobLevel
Income_JobLv <- ggplot(attrition,aes(x=as.factor(JobLevel),y=MonthlyIncome))+
    geom_boxplot()
Income_JobLv
```
proved our finding from the correlation matrix, JobLevel and MonthlyIncome are strongly positively correlated
Do employees staying at the company longer always tend to stay?
```{r}
#StayingPercentage wrt YearsAtCompany
stay <- table(attrition$Attrition,attrition$YearsAtCompany)[1,]
total <- table(attrition$Attrition,attrition$YearsAtCompany)[1,]+table(attrition$Attrition,attrition$YearsAtCompany)[2,]
staypct <- round(stay/total,3)
staypct
```
Employees staying longer in the company tend to stay with the company more comparing to new employees
```{r}
#density plot of YearsAtCompany vs Attrition
YearInComp_attri <- ggplot(attrition)+
  geom_density(aes(x=YearsAtCompany,fill=Attrition),alpha=0.8)
YearInComp_attri
```
overtime & businesstravel can naively tell degree of tiredness
use box-plot to see how overtime and travel influence salary
```{r}
tiredness_income <- ggplot(attrition,aes(x=OverTime,y=MonthlyIncome))+
                    geom_boxplot()+
                    facet_grid(~BusinessTravel)
tiredness_income
```

*arules mining preparation*
```{r}
#convert Likert Scales columns to factors
str(attrition)
conv_col <- c("Education","EnvironmentSatisfaction","JobInvolvement","JobLevel","JobSatisfaction","PerformanceRating","RelationshipSatisfaction","StockOptionLevel","WorkLifeBalance")
attrition[,conv_col] <- lapply((attrition[,conv_col]),as.factor)
```

```{r}
#convert numerical columns to factors for arules mining
#check distribution, then discretize accordingly
#example with how to discretize Age
hist(attrition$Age,breaks=10)
#found it is normally distributed
#try different discretization methods and breaks, such as cluster,frequency,fixed, see which works the best and fits the distribution
attrition$Age <- arules::discretize(attrition$Age,method = "cluster",breaks = 3, labels = c(1,2,3))
#do the same thing for other attributes
attrition$DailyRate <- arules::discretize(attrition$DailyRate,method = "cluster",breaks = 3, labels = c("low","mid","high"))
attrition$DistanceFromHome <- arules::discretize(attrition$DistanceFromHome,method = "cluster",breaks = 3, labels = c("low","mid","high"))
attrition$HourlyRate <- arules::discretize(attrition$HourlyRate,method = "cluster",breaks = 3, labels = c("low","mid","high"))
#distribution of monthlyIncome spreads out, so give more breaks
attrition$MonthlyIncome <- arules::discretize(attrition$MonthlyIncome,method = "cluster",breaks = 5, labels = c(1,2,3,4,5))
attrition$MonthlyRate <- arules::discretize(attrition$MonthlyRate,method = "cluster",breaks = 3, labels = c("low","mid","high"))
attrition$NumCompaniesWorked <- arules::discretize(attrition$NumCompaniesWorked,method = "cluster",breaks = 3, labels = c("few","several","many"))
#percentsalaryhike right skewed
attrition$PercentSalaryHike <- arules::discretize(attrition$PercentSalaryHike,method = "cluster",breaks = 4, labels = c(1,2,3,4))
attrition$TotalWorkingYears <- arules::discretize(attrition$TotalWorkingYears,method = "cluster",breaks = 4, labels = c(1,2,3,4))
#years in current role distribution has two local peaks, frequency&cluster do not work well, assign breaks manually
attrition$YearsInCurrentRole <- cut(attrition$YearsInCurrentRole,breaks = c(0,2.1,6.1,9.1,100),labels = c(1,2,3,4),right = FALSE)
attrition$TrainingTimesLastYear <- arules::discretize(attrition$TrainingTimesLastYear,method = "cluster",breaks = 3, labels = c("few","several","many"))
attrition$YearsAtCompany <- arules::discretize(attrition$YearsAtCompany,method = "cluster",breaks = 3, labels = c("few","several","many"))
attrition$YearsSinceLastPromotion <- arules::discretize(attrition$YearsSinceLastPromotion,method = "cluster",breaks = 3, labels = c("recently","a while","long ago"))
#years with current manager distribution has two local peaks, frequency&cluster do not work well, assign breaks manually
attrition$YearsWithCurrManager <- cut(attrition$YearsWithCurrManager,breaks = c(0,2.1,6.1,9.1,100),labels = c(1,2,3,4),right = FALSE)
str(attrition)
```
all the attributes are converted to factors now

*Association Rules Mining*
apply association rules with default hyper-parameters
```{r}
rules <- apriori(data = attrition,appearance = list(rhs = "Attrition=No",default ='lhs'))
inspect(head(sort(rules,by="lift",decreasing = T),10))
```

{}found on lhs, redundant rules found

adjust hyper-parameters and try use association rules again
```{r}
rules <- apriori(data = attrition,parameter = list(support = 0.05,confidence=0.8,minlen = 2),appearance = list(rhs = "Attrition=No",default ='lhs'))
inspect(head(sort(rules,by = "lift",decreasing = T),10))
```
rules with top lift values are too complex i.e. including many predictors, may overfit the data

```{r}
rules <- apriori(data = attrition,parameter = list(support = 0.3,confidence=0.9,maxlen = 4),appearance = list(rhs = "Attrition=No",default ='lhs'))
inspect(head(sort(rules,by = "lift",decreasing = T),10))
```
rules are too rigid with fewer predictors. They are not capturing salient patterns leading to no attrition, can be proved by the decrease of lift

final result
```{r}
rules <- apriori(data = attrition,parameter = list(support = 0.04,confidence=0.6,minlen = 2,maxlen = 6),appearance = list(rhs = "Attrition=No",default ='lhs'))
#remove redundant rules
subset_rules <- which(colSums(is.subset(rules,rules)) > 1)
rules <- sort(rules[-subset_rules], by  = "lift", decreasing = T)
inspect(head(rules,5))
```
top five rules found leading to Attrition="No"
 JobRole = Research Director
 JobRole = Manager
 MonthlyIncome = 5 (highest Income)
 JobLevel = 4 (very high)
 YearsWithCurrentManager = 4 (stayed 10+ years with current manager)

plot the rules 
```{r}
plot(rules,measure = c("support","lift"),shading="confidence")
```

*do the same thing on Attrition=Yes*
run arules with default hyper-parameter
```{r}
rules <- apriori(data = attrition,appearance = list(rhs = "Attrition=Yes",default ='lhs'))
inspect(head(sort(rules,by="lift",decreasing = T),10))
```
no rules found, because support=0.1 is too high, we only have 185 rows of Attrition="Yes"

tried different support values and found 0.013 is an appropriate choice for this dataset
```{r}
rules <- apriori(data = attrition,parameter = list(support = 0.013,confidence = 0.9,minlen = 2,maxlen = 6),appearance = list(rhs = "Attrition=Yes",default ='lhs'))
inspect(head(sort(rules,by = "lift",decreasing = T),10))
#remove redundant rules
subset_rules <- which(colSums(is.subset(rules,rules)) > 1)
rules <- sort(rules[-subset_rules], by  = "lift", decreasing = T)
inspect(head(rules,5))
```
top 5 rules by lift
Age=1,BusinessTravel=Travel_Frequently,JobLevel=1,MaritalStatus=Single
Age=1,BusinessTravel=Travel_Frequently,MaritalStatus=Single,MonthlyIncome=1
Education=3,JobLevel=1,OverTime=Yes,StockOptionLevel=0,YearsInCurrentRole=1
Age=1,JobInvolvement=3,OverTime=Yes,StockOptionLevel=0,YearsInCurrentRole=1
Age=1,BusinessTravel=Travel_Frequently,MonthlyIncome=1,StockOptionLevel=0

plot the rules 
```{r}
plot(rules,measure = c("support","lift"),shading="confidence")
```

*interpretation of results*
For Attrition="Yes":

Since the majority of the attrition attribute is "No" as employee stays, suport(Attrition=No) is high, and it is very hard to generate rules with very high lift.
Because the background is composed of mostly rows with "Attrition=Yes", the background confidence of achieving "Attrition=No" is
already high. Thus, our top five rules with highest lift values have very high confidence values. Looking at the top five rules obtained,
we have research director and manager as JobRole leading to employee staying at the company, and it is consistent with the thid and the fourth rule found, which
are JobLevel=4 and MonthlyIncome=5. Lastly, we have YearsWithCurrentManager=4, standing for working under the current manager for 10+ years. As we can expect,
a large portion of this group are on the manager/director level, which again, overlaps with employees identified by the first four rules. So in general, the well-paid
high-level managers/directors in this company tend to stay with the company comparing to the rest of the employees.


For Attrition="No":

Since we only have 185 rows with "Attrition=Yes", i.e. the employee has left the company, the support values of the top rules are very small. One obvious different
between the rules leading to attrition and stay is that the rules leading to attrition are more complicated. This can be explained by the correlation matrix and the 
distribution of Attrition. Since there is no single attribute or a specific level (with many counts) of any factor strongly correlated to "Attrition=Yes", we can not 
expect a single predictor as a rule with high lift leading to attrition. The top five rules of attrition all have lift values larger than 6, because support("Attrition=Yes) is very low, and confidence values of our top rules are pretty high. Obviously, some predictors appear multiple times among the top five rules, for example, Age=1 (<=32 yeras old), BusinessTravel=Travel_frequently, Years with Manger=1 (<= 2 years), and StockOptionLevel=0. These predictors are targeting the young entry-level employees with lower income and high pressure from business travels. 

Overall, the top rules leading to retention and attrition are consistent with our domain knowledge in the HR field. Young entry-level employees tend to job-hop for
higher income or lower pressure, while the decision makers tend to retain.

*shiny app for Association Rules Mining*
https://fyuan04.shinyapps.io/hw1Fangzhou/
